Synthesis of Network Intrusion Detection Rules

### 2019-10-16 paper/paper.tex
pass on intro

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first
% footnote. If that is unneeded, please comment it out.

\usepackage{booktabs}
\usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
\usepackage[caption=false, font=footnotesize]{subfig}
%%%%%%%%%%%%% code listing
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=false,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  %breaklines=true,            % Zeilen werden Umgebrochen
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

%% \usepackage{cite}
%% \usepackage{amsmath,amssymb,amsfonts}
%% \usepackage{algorithmic}
%% \usepackage{graphicx}
%% \usepackage{textcomp}
%% \usepackage{xcolor}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Search-based Synthesis of Network Intrusion Detection Rules}

%%%%%%%%%%% Anonymized

%% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address}
%% \and
%% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address}
%% \and
%% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address}
%% \and
%% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address}
%% \and
%% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address}
%% \and
%% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address}
%% }

\maketitle

\begin{abstract}
...
\end{abstract}

\begin{IEEEkeywords}
NIDS, synthesis, search
\end{IEEEkeywords}

\section{Introduction}

Network Intrusion Detection Systems (\nids{}) are software systems
that monitor the network traffic for malicious behavior and act
accordingly by blocking messages or alerting humans about suspicious
events~\cite{Mitchell:2014:SID:2597757.2542049}. \nids{} are typically
placed behind a firewall, vetting the traffic that the firewall did
not block. Various open-source (\eg{}, Snort~\cite{snort} and
Suricata~\cite{suricata}) and commercial implementations (\eg{},
SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar}) exist
today. These systems are very popular in industry to secure local
computer networks given the amount of potential malicious traffic that
exist on the Internet.

\sloppy \nids{} are typically categorized in two
groups~\cite{kumar2007survey}: rule-based and anomaly-based \nids. A
rule-based intrusion detector\footnote{\aka\ signature-based intrusion
  detector.} checks if the network traffic matches a fixed set of
rules. Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
prescribes a method to capture a denial-of-service
attack~\cite{understanding-dos} to a server by matching specific
conditions about the network traffic. Relevant conditions appear in
bold. Rule-based \nids{} focus on known attacks whereas anomaly-based
\nids{} focus on unknown attacks that could be observed by first
learning the regular traffic and then capturing suspicious
traffic~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}.

These two kinds of intrusion detection mechanisms are
complementary. Conceptually, rule-based NIDS could miss attacks (as
rulesets can become outdated) whereas anomaly-based NIDS can raise
false alarms (as the detection algorithms use are approximate).
Rule-based NIDS are restricted to a fixed set of rules and attackers
constantly create new strategies. Furthermore, manually writing rules
to capture those attacks is tedious and error-prone. As consequence,
IT-security companies advertize paid rulesets for consumers to
buy~\cite{proofpoint-etpro,snort-rule-subscriptions}.  Despite this
fundamental issue, rule-based systems are tremendously popular in
industry today~\Fix{cite}. Anomaly-based NIDS, in contrast, are not
restricted to a fixed set of rules. They learn usage behavior from
positive traffic and, based on that, alert uncommon
behavior~\cite{7579764}. Unfortunately, anomaly-based NIDS are
fundamentally imprecise. To sum up, both approaches have positive and
negative aspects.

@@

This paper proposes a technique, dubbed \tname{}, to close the gap
between these two strategies. \tname{} automatically synthesizes rules
from positive and negative examples (\ie{}, benign and malicious
traffic). A scenario of application of \tname{} is one where an
anomaly-based NIDS identifies abnormal traffic and signals that
traffic to \tname{} to craft rules that can be used by rule-based
NIDS. Note that anomaly-based NIDS are evolving pretty quick with the
advances in machine learning, but rule-based NIDS are still extremely
popular.  \tname{} could also leverage existing databases of malicious
traffic to synthesize rules.  Regardless of how the negative traffic
is produced (out of scope of this paper), those rules can be
distributed to rule databases for free. Our goal is to harden the
protection of NIDS by automatically synthesizing rules from negative
examples.

%%  The
%% core motivation is that 1) attackers are productive in creating new
%% ways to circumvent existing protections and 2) manual creation of
%% rules is tedious and time-consuming.

@@

\begin{figure*}[t]
\centering
\includegraphics[trim=10 360 182 130,clip,width=0.7\textwidth]{figs/nids-workflow}
\caption{\tname\ workflow.}
\label{fig:overview}
\end{figure*}

@@

\balance
%\vspace*{-2\baselineskip}
\bibliographystyle{IEEEtran}
\bibliography{references}

@@

%%  LocalWords:  Suricata SolarWinds QRadar NIDS OISF TCP ACK msg tcp
%%  LocalWords:  encodings eval cardinality

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
\usepackage[caption=false, font=footnotesize]{subfig}
%%%%%%%%%%%%% code listing
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=false,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  %breaklines=true,            % Zeilen werden Umgebrochen
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Synthesis of Network Intrusion Detection Rules}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Alcantara\inst{1}\Comment{\orcidID{0000-1111-2222-3333}} \and
Guilherme Padilha\inst{1}\Comment{\orcidID{1111-2222-3333-4444}} \and
Marcelo d'Amorim\inst{1}\Comment{\orcidID{0000-0002-1323-8769}}}
%
\authorrunning{Alcantara et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
...
\keywords{NIDS, synthesis, search}
\end{abstract}
%
%
%
\section{Introduction}

Network Intrusion Detection Systems (\nids{}) are software systems
that monitor the network traffic for malicious behavior and act
accordingly by blocking messages or alerting humans about suspicious
events~\cite{Mitchell:2014:SID:2597757.2542049}. \nids{} are typically
placed behind a firewall, vetting the traffic that the firewall did
not block. Various open-source (\eg{}, Snort~\cite{snort} and
Suricata~\cite{suricata}) and commercial NIDS implementations (\eg{},
SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar}) exist
today. These systems are very popular in industry to secure local
computer networks given the amount of potential malicious traffic that
exist on the Internet.

\sloppy \nids{} are typically categorized in two
groups~\cite{kumar2007survey}: rule-based and anomaly-based \nids. A
rule-based intrusion detector\footnote{\aka\ signature-based intrusion
  detector.} checks if the network traffic matches a fixed set of
rules. Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
prescribes a method to capture a denial-of-service
attack~\cite{understanding-dos} to a server by matching specific
conditions about the network traffic. Relevant properties about the
traffic of interest appear in bold (see
Section~\ref{sec:suri-metas-coverage}). Rule-based \nids{} focus on
known attacks whereas anomaly-based \nids{} focus on unknown attacks
that could be observed by first learning the regular traffic and then
acting
properly~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}. These
two kinds of intrusion detection mechanisms are
complementary. Rule-based NIDS could miss attacks (as rulesets can
become outdated) whereas anomaly-based NIDS can raise false alarms (as
detection algorithms are approximate).

Rule-based NIDS are restricted to a fixed set of rules defined by the
network system administrator. Manually writing these rules is tedious,
error-prone, but important as attackers constantly create new
strategies to circumvent existing rules. IT-security companies
capitalize on this phenomena and put rulesets on the market for
sale~\cite{proofpoint-etpro,snort-rule-subscriptions}.  Despite this
fundamental issue, rule-based systems are tremendously popular in
industry today. Anomaly-based NIDS are not restricted to a fixed set
of rules. They learn usage behavior from positive traffic and, based
on that, alert uncommon behavior~\cite{7579764}. Unfortunately,
anomaly-based NIDS are fundamentally imprecise. To sum up, rule-based
NIDS are precise but incomplete whereas anomaly-based NIDS are (more)
complete but imprecise.

@@

This paper proposes \tname{}, a technique that uses machine
intelligence to automatically synthesize rules for rule-based
\nids. Our general goal is to harden the protection of NIDS by
improving the creation process of rules. One scenario of application
of \tname{} is one where an anomaly-based NIDS identifies abnormal
traffic and signals that traffic to \tname{} to create and distribute
the rules.

%% Note that anomaly-based NIDS are evolving pretty quick with the
%% advances in machine learning, but rule-based NIDS are still extremely
%% popular.  \tname{} could also leverage existing databases of malicious
%% traffic to synthesize rules.  Regardless of how the negative traffic
%% is produced (out of scope of this paper), those rules can be
%% distributed to rule databases for free.
%%  The
%% core motivation is that 1) attackers are productive in creating new
%% ways to circumvent existing protections and 2) manual creation of
%% rules is tedious and time-consuming.

\tname{} synthesizes rules from positive and negative examples, \ie{},
malicious and benign traffic.

@@

\begin{figure*}[t]
\centering
\includegraphics[trim=10 360 182 130,clip,width=0.95\textwidth]{figs/nids-workflow}
\caption{\tname\ workflow.}
\label{fig:overview}
\end{figure*}

@@

\Fix{...}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{references}
%

@@
